Biden Forcested To Win the Election

### 2020-11-03 outputs/paper/paper2.Rmd
Adding final version

---
title: "Biden Forcested To Win the Election"
subtitle: "Based on US Demographics of Age, Education, Household Income and Race"
author: "Arjun Dhatt, Benjamin Draskovic, Yiqu Ding, Gantavya Gupta"
thanks: "Code and data are available at: https://github.com/STA304-PS4/Trump-vs-Biden."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: |
  | In this report, we will be forecasting the winner of the upcoming US Presidential Elections 2020 between incumbent Donald Trump of the Republican Party and Joe Biden of the Democratic Party, by looking at different factors that affect a citizen’s voting intentions leading to Election Day. We cleaned the Nationscape dataset and chose the variables age, education, race, family income, geographic region, and mask-wearing decision which will form the basis of our Elections prediction. Then we used an MRP model (a combination of multilevel logistic regression and post-stratification) to analyse the factors and it also helps us in achieving a much narrower confidence interval for our prediction. After our analysis, we concluded that Joe Biden will win the overall popular vote of the 2020 American presidential election.  
output:
  bookdown::pdf_document2:
    fig_caption: yes        
    includes:  
      in_header: modify.tex
toc: FALSE
bibliography: references.bib
header-includes:
- \usepackage{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(dplyr)
library(lme4)
library(merTools)
library(readr)
library(knitr)
```

@@

The variables that we used in our dataset are as follow: 
-	age
     o	“What is your age”
     o	It provides an integer value greater than or equal to 18
-	education
     o	“What is the highest level of education you have completed?” 
     o	An extensive list of options are provided
-	Hispanic
     o	“Are you of Hispanic, Latino, or Spanish origin?” 
     o	An extensive list of options are provided 
-	household_income
     o	“What is your current annual household income before taxes?" 
     o	Provides an extensive list of options beginning with $5,000 increments and then increases to $25,000 increments
     o	This is the variable in our dataset that many people did not feel comfortable answering and was left as N/A. When        cleaning the dataset, we removed all responses that left this answer as N/A  
-	race_ethnicity
     o	“What is your race”
     o	An extensive list of options are provided. 
-	extra_covid_worn_mask
     o	“Have you done any of the following in the past week? - Worn a mask when going out in public” 
     o	Yes or No options are available for the respondent to fill out 
-	state
     o	“Respondent’s state, as a two-character postal abbreviation. Calculated based on entered ZIP code.” 
     o	Manually filled out by respondent 
-	trump_biden
     o	“If the general election for president of the United States was a contest between Joe Biden and Donald Trump, who        would you support? “ 
     o	Options are either ‘Joe Biden’, ‘Donald Trump’, or ‘Don’t Know’

We created an additional variable called ‘binary’ which outputs the result of the variable ‘trump_biden’ as a 1 or a 0 making it binary, therefore easier to analyze. Since some variables had an extensive list of options (e.g. age, household income), we grouped the responses for those variables so that they are simpler and easier to analyze.

The first few rows of the cleaned dataset can be viewed below: 
```{r}
tbdata <- read.csv("clean-survey-2.csv")
tbdatashort <- tbdata %>% 
  dplyr::select(
    age_group, race_ethnicity, household_income_grouped, trump_biden
  )
Table1 <- tbdatashort %>% 
  head() %>% 
  kable(align = "c",
        format = "simple",
        col.names = c("Age Group", "Race", "Household Income", "Presidential choice")
        ) %>% 
  kable_styling(font_size = 5)
Table1
```

In Figure [] below, we can see that the distribution is largely skewed towards families with income less than $50,000:
```{r}
# Setup data for the pie chart and mutate a variable that details proportions for various responses. 
incomeinfo <- tbdata %>% group_by(household_income_grouped) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
# Graph the pie chart using ggplot; Set colour scheme, font, etc. 
incomegraph <- incomeinfo %>% ggplot(aes(x = "", y = prop, fill = household_income_grouped)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Accent") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "Survey Distribution of Household Income", fill="Household Income") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4.7) +
     theme(text = element_text(size=15))
incomegraph
```
Because of the skewed distribution of income, we can justify using post stratification. Clearly, those with income less than $50,000 are overrepresented in the dataset as they take up 50% of the responses, therefore it is appropriate to use post stratification.

We can see a similar skewed distribution in Figure [] below: 
```{r}
# Setup data for the pie chart and mutate a variable that details proportions for various responses. 
raceinfo <- tbdata %>% group_by(race_ethnicity) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
# Graph the pie chart using ggplot; Set colour scheme, font, etc. 
racegraph <- raceinfo %>% ggplot(aes(x = "", y = prop, fill = race_ethnicity)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Set3") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "Survey Distribution of Respondent Race", fill="What is your race?") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4) +
     theme(text = element_text(size=13))
racegraph
```
Looking at the distribution of race, we can see that 75% of the respondants from the survey are white. Since the distribution does not represent minorities, we can justify using post stratification.

Lastly, in Figure [] below, we can see how the population aged older than 76 are barely represented: 
```{r}
# Setup data for the pie chart and mutate a variable that details proportions for various responses. 
ageinfo <- tbdata %>% group_by(age_group) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
# Graph the pie chart using ggplot; Set colour scheme, font, etc. 
agegraph <- ageinfo %>% ggplot(aes(x = "", y = prop, fill = age_group)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Set3") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "Survey Distribution of Respondent Age", fill="Age Groups") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4)+
     theme(text = element_text(size=13))
agegraph
```
The fact that those age 76+ make up only a 3% of the data could be a product of the fact that the survey was completed online. Clearly, the survey doesn't accurately represent those aged older than 76 posing the question — does our data represent the United States? The answer is no. To adjust for the fact that the population aged 76+, we will use post stratification.

The dataset that we used contains many key features. Many of the variables in the dataset contain an extensive list of options assuring us that the data represents the participant's view accurately. Despite the fact that we cleaned the dataset to make it simpler to view, we still believe that the detailed responses that the participants provided accurately reflect their views on the 2020 election, making our analysis meaningful. Another key feature is the fact that the data was collected over a short period of time— from June 25th, 2020 to July 1st, 2020. This significantly reduces the negative time based effects that may have affected the data ensuring that all respondents are responding on a similar ground and no one's opinion dramatically differed due to the survey taking place over a large period of time.

The dataset we used contains many strengths and weaknesses. One of the strengths is due to the fact that the survey is distributed online. In today’s age where most things are online, this ensures that we can reach a wide part of the United States. At the same time, this can also be viewed as a weakness because certain age groups of the United States may not be tech-savvy enough to complete an online survey. The older population of the united states (e.g. 75+) may not be able to complete the survey because it is online which means that our data may not accurately reflect the older populations views regarding the 2020 election. A potential solution to this weakness is to use administer multiple forms of the survey to ensure that all people can access the survey.

Another weakness of the data comes from the fact that we use convenience-based sampling which inherently contains bias. Because the sampling method is based on convenience, we may be unlucky when sampling and get a sample group that inaccurately represents the population of the US making our results meaningless.

Lastly another weakness of our dataset is due to the fact that the survey data we are analyzing is collected from June 25th, 2020 to July 1st, 2020. Between July and November (the month of the election), many voters may not have an informed vote and their opinions may still be swayed after watching crucial debates and hearing what the leaders have to say; our data doesn’t take into consideration the people whose votes may change in the most crucial months of the election. A solution to this weakness can be to analyze data that may have occurred closer to the actual election so that the participants have an informed vote.

@@

### Summary Table
\begin{center}
\captionof{table}{statistics summary for logistic regression}
\begin{tabular}{|l|l|l|l|}
\hline
variable & $\hat{\beta}$ & Std. Error & p-value \\ \hline
intercept & -0.00 & 0.26 & 0.99 \\ \hline
$36 < age < 55$ & $0.42$ & $0.07$ & $6.06e^{-9}$ \\ \hline
$56 < age < 75$ & $0.37$ & $0.07$ & $8.91e^{-7}$ \\ \hline
$age > 76$ & $0.62$ & $0.18$ & $0.00$ \\ \hline
Asian or Pacific Islander & $ -0.75 $ & $0.27$ & $0.01$ \\ \hline
black & $-1.89$ & $0.26$ & $1.25e^{-12}$ \\ \hline
nec & $-0.75$ & $0.26$ & $0.01$ \\ \hline
white & $0.09$ & $0.24$ & $0.68$ \\ \hline
$income > 150k$ & $0.14$ & $0.11$ & $0.22$ \\ \hline
$income < 25k$ & $-0.38$ & $0.10$ & $0.01$ \\ \hline
$25k < income < 50k$ & $-0.27$ & $0.10$ & $0.01$ \\ \hline
$50k < income < 75k$ & $-0.19$ & $0.10$ & $0.06$ \\ \hline
$75k < income < 100k$ & $-0.31$ & $0.11$ & $0.01$ \\ \hline
\end{tabular}
\end{center}

@@

```{r}
# Read in the csv file for post stratification
post_strat_cellcount_2 <- read_csv("post_strat_cellcount-2.csv")
# get post stratification estimates using predict()
post_strat_cellcount_2$estimate <-
  mod %>%
  predict(newdata = post_strat_cellcount_2, allow.new.level = TRUE, type="response")
```

```{r, fig.cap='95% Prediction Intervals for forcasting the Chances that Donald Trump wins the election'}
PI <- predictInterval(merMod = mod, newdata = post_strat_cellcount_2,                         
                      level = 0.95, n.sims = 1000,                         
                      stat = "median", type= "probability",                      
                      include.resid.var = TRUE)

ggplot(aes(x=1:30, y=fit, ymin=lwr, ymax=upr), data=PI[1:30,]) +   
  geom_point() +   geom_linerange() +   
  labs(x="Index", y="Prediction w/ 95% PI") + 
  theme_bw()
```

@@

Looking at the income break down of our data it is clear when comparing Figure 1 and Figure 4 that the surveyed population tended towards a lower income on average. This is an interesting occurrence and may have to do with the tech literacy having a larger survey biasing effect than literacy and tech access. On average older individuals have higher incomes so it could be that the data is skewing younger and that is what is causing this. 
## Figure 1
```{r}
incomegraph
```

## Figure 4
```{r}
acs <- read_dta("post_strat_data-2.dta")
acs <- labelled::to_factor(acs)
incomeinfoacs <- acs %>% group_by(household_income_grouped) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
incomegraphacs <- incomeinfoacs %>% ggplot(aes(x = "", y = prop, fill = household_income_grouped)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Accent") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "ACS Distribution of Household Income", fill="Household Income") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4.7) +
     theme(text = element_text(size=15))

@@

## Figure 2
```{r}
agegraph
```

## Figure 5
```{r}
ageinfoacs <- acs %>% group_by(age_group) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
agegraphacs <- ageinfoacs %>% ggplot(aes(x = "", y = prop, fill = age_group)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Accent") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "ACS Distribution of Respondent Age", fill="Household Income") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4.7) +
     theme(text = element_text(size=15))

@@

## Figure 3
```{r}
racegraph
```

## Figure 6
```{r}
raceinfoacs <- acs %>% group_by(race_ethnicity) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
racegraphacs <- raceinfoacs %>% ggplot(aes(x = "", y = prop, fill = race_ethnicity)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Accent") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "ACS Distribution of Respondent Race", fill="Household Income") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4.7) +
     theme(text = element_text(size=15))

@@

## Table 6
```{r}
Table6 <- post_strat_cellcount_2 %>% 
  dplyr::select(
    age_group, race_ethnicity, household_income_grouped, n, prop
  ) %>% 
  head() %>% 
  kable(align = "c",
        format = "simple",
        col.names = c("Age Group", "Race", "Household Income", "Number of Repondents", "Proportion of Responses"),
        digits = 6
        ) %>% 
  kable_styling(font_size = 5)
Table6
```

@@

```{r include=FALSE}
aggregate(tbdata[8], list(tbdata$race_ethnicity), mean)

@@

>>>>

---
title: "Biden Forcested To Win the Election"
subtitle: "Based on US Demographics of Age, Education, Household Income and Race"
author: "Arjun Dhatt, Benjamin Draskovic, Yiqu Ding, Gantavya Gupta"
thanks: "Code and data are available at: https://github.com/STA304-PS4/Trump-vs-Biden."
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::pdf_document2:
    fig_caption: yes        
    includes:  
      in_header: modify.tex
toc: FALSE
bibliography: references.bib
header-includes:
- \usepackage{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(dplyr)
library(lme4)
library(merTools)
library(readr)
library(knitr)
library(haven)
```
# Abstract

In this report, we will be forecasting the winner of the upcoming US Presidential Elections 2020 between incumbent Donald Trump of the Republican Party and Joe Biden of the Democratic Party, by looking at different factors that affect a citizen’s voting intentions leading to Election Day. We cleaned the Nationscape dataset and closely examine the variables age, race, household income, and vote which will form the basis of our election prediction. Then we used an MRP model (a combination of multilevel logistic regression and post-stratification) to analyse the factors and it also helps us in achieving a much narrower confidence interval for our prediction. After our analysis, we concluded that Joe Biden will win the overall popular vote of the 2020 American presidential election.

# Keywords
- Forecasting
- US 2020 Election
- Multi-level Regression with Post-Stratification
- Trump
- Biden

@@

The variables that we used in our dataset are as follow:

- age
     o	Asks the question “What is your age?”
     o	It provides an integer value greater than or equal to 18 
- household_income
     o	Asks the question “What is your current annual household income before taxes?" 
     o	Provides an extensive list of options beginning with $5,000 increments and then increases to $25,000 increments. We         later grouped the options into larger bins making it simpler to understand and easier to analyze
     o	This is the variable in our dataset that many people did not feel comfortable answering and was left as N/A. When           cleaning the dataset, we removed all responses that left this answer as N/A  
- race_ethnicity
     o	Asks the question “What is your race?”
     o	An extensive list of options are provided for the respondent to disclose their race. Because the list is very               extensive, we decided to group the options into larger bins making to simpler to understand and easier to analyze. 
- trump_biden
     o	Asks the question “If the general election for president of the United States was a contest between Joe Biden and           Donald Trump, who would you support? “ 
     o	Options are either ‘Joe Biden’, ‘Donald Trump’, or ‘Don’t Know’

We created an additional variable called ‘binary’ which outputs the result of the variable ‘trump_biden’ as a 1 or a 0 making it binary, therefore easier to analyze. Since some variables had an extensive list of options (e.g. age, household income, race), we grouped the responses for those variables so that they are simpler and easier to analyze.

To develop a sense of familiarity with the dataset, we will closely examine some of the variables we will be analyzing in the dataset.

The first few rows of the cleaned dataset can be viewed in Table 1 below:
### Table 1 
```{r, tidy =FALSE, echo =FALSE, warning= FALSE}
# Read in the data 
tbdata <- read.csv("clean-survey-2.csv")
# Create shortened data set  
tbdatashort <- tbdata %>% 
  dplyr::select(
    age_group, race_ethnicity, household_income_grouped, trump_biden
  )
Table1 <- tbdatashort %>% 
  head() %>% 
  kable(align = "c",
        format = "simple",
        col.names = c("Age Group", "Race", "Household Income", "Presidential choice")
        ) 
Table1
```

In Figure 1 below, we will examine the variable 'household_income' and we can see that our data is largely skewed towards households with income less than $50,000:
### Figure 1 
```{r, tidy =FALSE, echo =FALSE, warning= FALSE}
# Setup data for the pie chart and mutate a variable that details proportions for various responses. 
incomeinfo <- tbdata %>% group_by(household_income_grouped) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
# Graph the pie chart using ggplot; Set colour scheme, font, etc. 
incomegraph <- incomeinfo %>% ggplot(aes(x = "", y = prop, fill = household_income_grouped)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Accent") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "Survey Distribution of Household Income", fill="Household Income") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4.7) +
     theme(text = element_text(size=15))
incomegraph
```
Clearly, those with income less than $50,000 are overrepresented in the dataset as they take up 50% of the responses, therefore it is appropriate to use post stratification on our dataset.

We can see a similar skewed distribution with variable 'race' in Figure 2 below:
### Figure 2
```{r, tidy =FALSE, echo =FALSE, warning= FALSE}
# Setup data for the pie chart and mutate a variable that details proportions for various responses. 
raceinfo <- tbdata %>% group_by(race_ethnicity) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
# Graph the pie chart using ggplot; Set colour scheme, font, etc. 
racegraph <- raceinfo %>% ggplot(aes(x = "", y = prop, fill = race_ethnicity)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Set3") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "Survey Distribution of Respondent Race", fill="What is your race?") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4) +
     theme(text = element_text(size=13))
racegraph
```
It is obvious that the respondents of our survey are primarily white and there is very little representation of minorities.

Lastly, by examining the variable age in Figure 3 below, we can see how very little the population 76+ is represented in our dataset. 
### Figure 3
```{r, tidy =FALSE, echo =FALSE, warning= FALSE}
# Setup data for the pie chart and mutate a variable that details proportions for various responses. 
ageinfo <- tbdata %>% group_by(age_group) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
# Graph the pie chart using ggplot; Set colour scheme, font, etc. 
agegraph <- ageinfo %>% ggplot(aes(x = "", y = prop, fill = age_group)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Set3") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "Survey Distribution of Respondent Age", fill="Age Groups") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4)+
     theme(text = element_text(size=13))
agegraph
```
The fact that those age 76+ make up only 3% of the data could be a product of the fact that the survey was completed online. Clearly, the survey doesn't accurately represent those aged older than 76/ This poses the question — does our data represent the United States?  By looking at the figure above, the answer is no. To adjust for the fact that the population aged 76+, we will use post stratification.

The dataset that we used contains many key features. Many of the variables in the dataset contain an extensive list of options assuring us that the data represents the participant's view accurately. Despite the fact that we cleaned the dataset to make it simpler to view, we still believe that the detailed responses that the participants provided accurately reflect their views on the 2020 election, making our analysis meaningful. Another key feature is the fact that the data was collected over a short period of time— from June 25th, 2020 to July 1st, 2020. This significantly reduces the negative time-based effects that may have affected the data ensuring that all respondents are responding on a similar ground and no one's opinion dramatically differed due to the survey taking place over a large period of time.

The dataset we used contains many strengths and weaknesses. One of the strengths is due to the fact that the survey is distributed online. In today’s age where most things are online, this ensures that we can reach a wide part of the United States. At the same time, this can also be viewed as a weakness because certain age groups of the United States may not be tech-savvy enough to complete an online survey. The older population of the United States (e.g. 75+) may not be able to complete the survey because it is online which means that our data may not accurately reflect the older populations views regarding the 2020 election. A potential solution to this weakness is to use administer multiple forms of the survey to ensure that all people can access the survey.

Another weakness of the data comes from the fact that we use convenience-based sampling which inherently contains bias. Because the sampling method is based on convenience, we may be unlucky when sampling and get a sample group that inaccurately represents the population of the US making our results meaningless. A potential solution to this is to use a random sampling-based method to reduce bias as much as possible.

Lastly another weakness of our dataset is due to the fact that the survey data we are analyzing is collected from June 25th, 2020 to July 1st, 2020. Between July and November (the month of the election), many voters may not have an informed vote and their opinions may still be swayed after watching crucial debates and hearing what the leaders have to say; our data doesn’t take into consideration the people whose votes may change in the most crucial months of the election. A solution to this weakness can be to analyze data that may have occurred closer to the actual election (e.g. September or October) so that the participants have an informed vote.

@@

### Table 2
### Summary Table
\begin{center}
\captionof{table}{statistics summary for logistic regression}
\begin{tabular}{|l|l|l|l|}
\hline
variable & $\hat{\beta}$ & Std. Error & p-value \\ \hline
intercept & -0.00 & 0.26 & 0.99 \\ \hline
$36 < age < 55$ & $0.42$ & $0.07$ & $6.06e^{-9}$ \\ \hline
$56 < age < 75$ & $0.37$ & $0.07$ & $8.91e^{-7}$ \\ \hline
$age > 76$ & $0.62$ & $0.18$ & $0.00$ \\ \hline
Asian or Pacific Islander & $ -0.75 $ & $0.27$ & $0.01$ \\ \hline
black & $-1.89$ & $0.26$ & $1.25e^{-12}$ \\ \hline
nec & $-0.75$ & $0.26$ & $0.01$ \\ \hline
white & $0.09$ & $0.24$ & $0.68$ \\ \hline
$income > 150k$ & $0.14$ & $0.11$ & $0.22$ \\ \hline
$income < 25k$ & $-0.38$ & $0.10$ & $0.01$ \\ \hline
$25k < income < 50k$ & $-0.27$ & $0.10$ & $0.01$ \\ \hline
$50k < income < 75k$ & $-0.19$ & $0.10$ & $0.06$ \\ \hline
$75k < income < 100k$ & $-0.31$ & $0.11$ & $0.01$ \\ \hline
\end{tabular}
\end{center}

@@

@@

Looking at the income break down of our data it is clear when comparing Figure 1 and Figure 4 that the surveyed population tended towards a lower income on average. This is an interesting occurrence and may have to do with the tech literacy having a larger survey biasing effect than literacy and tech access. On average older individuals have higher incomes so it could be that the data is skewing younger and that is what is causing this. 
### Figure 1
```{r, tidy =FALSE, echo =FALSE, warning= FALSE}
incomegraph
```

### Figure 4
```{r}
# read in data
acs <- read_dta("post_strat_data-2.dta")
# label the data as it is a dta
acs <- labelled::to_factor(acs)
#create a tibble that includes counts and proportions based on groups
incomeinfoacs <- acs %>% group_by(household_income_grouped) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
#Graph this tibble in a pie graph
incomegraphacs <- incomeinfoacs %>% ggplot(aes(x = "", y = prop, fill = household_income_grouped)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Accent") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "ACS Distribution of Household Income", fill="Household Income") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4.7) +
     theme(text = element_text(size=15))

@@

### Figure 3
```{r, tidy =FALSE, echo =FALSE, warning= FALSE}
agegraph
```

### Figure 5
```{r}
#create a tibble that includes counts and proportions based on groups
ageinfoacs <- acs %>% group_by(age_group) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
#Graph this tibble
agegraphacs <- ageinfoacs %>% ggplot(aes(x = "", y = prop, fill = age_group)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Accent") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "ACS Distribution of Respondent Age", fill="Household Income") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4.7) +
     theme(text = element_text(size=15))

@@

### Figure 2
```{r, tidy =FALSE, echo =FALSE, warning= FALSE}
racegraph
```

### Figure 6
```{r, tidy =FALSE, echo =FALSE, warning= FALSE}
#create a tibble that includes counts and proportions based on groups
raceinfoacs <- acs %>% group_by(race_ethnicity) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
#Graph this tibble in a pie graph
racegraphacs <- raceinfoacs %>% ggplot(aes(x = "", y = prop, fill = race_ethnicity)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Accent") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "ACS Distribution of Respondent Race", fill="Household Income") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4.7) +
     theme(text = element_text(size=15))

@@

## Table 3
```{r, tidy =FALSE, echo =FALSE, warning= FALSE}
# Read in the csv file for post stratification
post_strat_cellcount_2 <- read_csv("post_strat_cellcount-2.csv")
#Generate a table that displays the first couple responses from the cell count that was generated for weighting 
Table6 <- post_strat_cellcount_2 %>% 
  dplyr::select(
    age_group, race_ethnicity, household_income_grouped, n, prop
  ) %>% 
  head() %>% 
  kable(align = "c",
        format = "simple",
        col.names = c("Age Group", "Race", "Household Income", "Number of Repondents", "Proportion of Responses"),
        digits = 6
        ) 
Table6
```

@@

```{r include=FALSE}
#Get general percentage of people voting for Trump by Race
aggregate(tbdata[8], list(tbdata$race_ethnicity), mean)
#get overal percentage of people voting for Trump
mean(tbdata$binary)  
```

@@

## Git Repository

The link to the Git Repository is:

## Code

Below you will also find the R code used in this analysis:

### Preamble- Setting Up Libraries and Clean Data

```{r, echo=TRUE, results = "hide", warning= FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(dplyr)
library(lme4)
library(merTools)
library(readr)
library(knitr)
library(haven)
```

```{r, echo=TRUE, results = "hide", warning= FALSE}
clean_survey_2 <- read_csv("clean-survey-2.csv")
mod <- glmer(binary ~ age_group + (1|education_grouped) + race_ethnicity + household_income_grouped,
           data = clean_survey_2,
           family = binomial()
           )
```

```{r, echo=TRUE, results = "hide", warning= FALSE}
# Read in the csv file for post stratification
post_strat_cellcount_2 <- read_csv("post_strat_cellcount-2.csv")
# get post stratification estimates using predict()
post_strat_cellcount_2$estimate <-
  mod %>%
  predict(newdata = post_strat_cellcount_2, allow.new.level = TRUE, type="response")
```

```{r, echo=TRUE, results = "hide", warning= FALSE}
aggregate(tbdata[8], list(tbdata$race_ethnicity), mean)

### Code for tables and figures used in analysis

### Table 1 
```{r, echo=TRUE, results = "hide", warning= FALSE}
# Read in the data 
tbdata <- read.csv("clean-survey-2.csv")
# Create shortened data set  
tbdatashort <- tbdata %>% 
  dplyr::select(
    age_group, race_ethnicity, household_income_grouped, trump_biden
  )
Table1 <- tbdatashort %>% 
  head() %>% 
  kable(align = "c",
        format = "simple",
        col.names = c("Age Group", "Race", "Household Income", "Presidential choice")
        ) 
Table1
```

### Figure 1 
```{r, echo=TRUE, results = "hide", warning= FALSE}
# Setup data for the pie chart and mutate a variable that details proportions for various responses. 
incomeinfo <- tbdata %>% group_by(household_income_grouped) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
# Graph the pie chart using ggplot; Set colour scheme, font, etc. 
incomegraph <- incomeinfo %>% ggplot(aes(x = "", y = prop, fill = household_income_grouped)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Accent") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "Survey Distribution of Household Income", fill="Household Income") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4.7) +
     theme(text = element_text(size=15))
incomegraph
```

### Figure 2
```{r, echo=TRUE, results = "hide", warning= FALSE}
# Setup data for the pie chart and mutate a variable that details proportions for various responses. 
raceinfo <- tbdata %>% group_by(race_ethnicity) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
# Graph the pie chart using ggplot; Set colour scheme, font, etc. 
racegraph <- raceinfo %>% ggplot(aes(x = "", y = prop, fill = race_ethnicity)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Set3") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "Survey Distribution of Respondent Race", fill="What is your race?") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4) +
     theme(text = element_text(size=13))
racegraph
```

### Figure 3
```{r, echo=TRUE, results = "hide", warning= FALSE}
# Setup data for the pie chart and mutate a variable that details proportions for various responses. 
ageinfo <- tbdata %>% group_by(age_group) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
# Graph the pie chart using ggplot; Set colour scheme, font, etc. 
agegraph <- ageinfo %>% ggplot(aes(x = "", y = prop, fill = age_group)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Set3") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "Survey Distribution of Respondent Age", fill="Age Groups") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4)+
     theme(text = element_text(size=13))
agegraph
```

```{r echo=TRUE, results = "hide", warning= FALSE}
# Read in the csv file for post stratification
post_strat_cellcount_2 <- read_csv("post_strat_cellcount-2.csv")
# get post stratification estimates using predict()
post_strat_cellcount_2$estimate <-
  mod %>%
  predict(newdata = post_strat_cellcount_2, allow.new.level = TRUE, type="response")
```

# Weighting code that did not run 
```{r echo=TRUE, results = "hide", warning= FALSE}
post_strat_cellcount_2 %>%
  mutate(president_predict_prop = estimate*prop) %>%
  group_by(education_grouped) %>%
  summarise(president_predict = sum(president_predict_prop)) %>% 
  as.data.frame() %>% 
  knitr::kable()
```

```{r, echo=TRUE, results = "hide", warning= FALSE}
PI <- predictInterval(merMod = mod, newdata = post_strat_cellcount_2,                         
                      level = 0.95, n.sims = 1000,                         
                      stat = "median", type= "probability",                      
                      include.resid.var = TRUE)

ggplot(aes(x=1:30, y=fit, ymin=lwr, ymax=upr), data=PI[1:30,]) +   
  geom_point() +   geom_linerange() +   
  labs(x="Index", y="Prediction w/ 95% PI") + 
  theme_bw()
```
## Figure 4
```{r, echo=TRUE, results = "hide", warning= FALSE}
incomegraph
```

## Figure 5
```{r, echo=TRUE, results = "hide", warning= FALSE}
acs <- read_dta("post_strat_data-2.dta")
acs <- labelled::to_factor(acs)
incomeinfoacs <- acs %>% group_by(household_income_grouped) %>% summarise(n=n()) %>% mutate(prop = 100*(n/sum(n)))
incomegraphacs <- incomeinfoacs %>% ggplot(aes(x = "", y = prop, fill = household_income_grouped)) + 
      geom_bar(width = 1, stat = "identity") + 
      coord_polar("y", start = 0) +
      scale_fill_brewer(palette = "Accent") +
      theme_void() +
      ylab("") + xlab("") + 
     labs(title = "ACS Distribution of Household Income", fill="Household Income") +
     geom_text(aes(label = paste0(round(prop), "%")), position = position_stack(vjust = 0.5), angle=45, size = 4.7) +
     theme(text = element_text(size=15))

## Figure 6
```{r, echo=TRUE, results = "hide", warning= FALSE}
agegraph
```
